---
title: "Chi"
author: "Kamila Kolpashnikova"
output:  
      html_document:  
        keep_md: true 
---

# Pearson's chi-squared test and Interpretation


```{r, echo=FALSE, message=FALSE}
require(knitr)
statapath <- "C:/Program Files (x86)/Stata12/StataSE-64.exe"
opts_chunk$set(engine="stata", engine.path=statapath, comment="")
```

The data here is derived from the [American Time Use Survey](http://www.bls.gov/tus/). For the coding procedures, contact me via email.

## Step 1. Downloading the dataset
You need to download the ATUS_full3.dta file from my github to your working directory.

```{r, results='hide'}
use ATUS_full3.dta, clear
```



## Step 2. Choose nominal variables

In the datase, I will use two dummy variables. One which is sex of the respondent (Female; 1-'Female', 2-'Male'), and the other is whether the respondent has children under 5 years of age (Under5; 1-'has children under 5 y/o', 0 - otherwise).

```{r, echo=2:3}
use ATUS_full3.dta, clear
tab Female
tab Under5
```

## Step 3. Peform chi-squared test

You use an option `chi` with the `tabulate` command.

```{r, echo=2}
use ATUS_full3.dta, clear
tab Female Under5, chi 
```

## Step 4. Interpretation

Chi square is not a PRE measure (PRE - proportional reduction of error), so you cannot tell the strength of the association between the two variables just by looking at the chi. But by using chi-squared test and its significance we can say a bit about existence of the relation between the two variables.

Thus in the results above the chi square with 1 degree of freedom is 91.1296, and it's p-value is rounded to 0.000. If we choose the alpha-level 0.001 (you can also choose 0.05, 0.01, etc.), we can say that:

The p-value provides us with enough evidence to reject the null hypothesis of statistical independence.

We are 99.9% ($(1-\alpha-level)*100$) confident that we can reject the null hypothesis. 
OR 
There is only 0.1% ($\alpha-level*100$) chance that the relation between `dependent variable` and `independent variable` is due to chance.

Therefore, we conclude that:

The two variables are associated, i.e., they are not statistically independent.

## Alternative Step 3. Chi-squared test, non-significant results

Let's use the same test but comparing whether we have more Females oversampled on the weekends.

```{r, echo=2}
use ATUS_full3.dta, clear
tab Weekday Female, chi
```

## Step 4. Interpretation

In the output above, the chi square with 1 degree of freedom is 0.3311 and it's p-value is rounded to 0.565. If we choose the alpha-level 0.05 (or any other conventional level), we can say that:

The p-value does not provide us with enough evidence to reject the null hypothesis of statistical independence.

We don't have enough evidence to reject the null hypothesis at 0.05 level. 

Therefore, we conclude that:

We don't have enough evidence to suggest that these two variables are associated.