---
title: "gamma"
author: "Kamila Kolpashnikova"
output:  
      html_document:  
        keep_md: true 
---

# Gamma test


```{r, echo=FALSE, message=FALSE}
require(knitr)
statapath <- "C:/Program Files (x86)/Stata12/StataSE-64.exe"
opts_chunk$set(engine="stata", engine.path=statapath, comment="")
```

The data here is derived from the [American Time Use Survey](http://www.bls.gov/tus/). For the coding procedures, contact me via email.

## Step 1. Downloading the dataset
You need to download the ATUS_full3.dta file from my github to your working directory.

```{r, results='hide'}
use ATUS_full3.dta, clear
```



## Step 2. Choose ordinal variables

In the datase, I will use two ordinal variables. One which is education (EduInYears) and another one is age group (Age).

```{r, echo=2:3}
use ATUS_full3.dta, clear
sum EduInYears
sum Age
```

## Step 3. Peform gamma test

You use an option `gamma` with the `tabulate` command.

```{r, echo=2}
use ATUS_full3.dta, clear
tab EduInYears Age, gamma
```

In the output above, the gamma = 0.0359 and Stata also shows ASE (Asymptotic Standard Error) = 0.002. To calculate z-value you need to divide gamma by ASE.

```{r}
display 0.0359/0.002
```

Using a table for z-values, we can find that for z=17.95, p <0.001 (let's assume that this is our alpha-level).

## Step 4. Interpretation

Gamma is a PRE measure (PRE - proportional reduction of error), so you can tell the strength of the association between the two variables using gamma. 

The p-value provides us with evidence to reject the null hypothesis of statistical independence.

We are more than 99.9% ($(1-\alpha-level)*100$) confident that we can reject the null hypothesis. 
OR 
There is less than 0.1% ($\alpha-level*100$) chance that the relation between `dependent variable` and `independent variable` is due to chance.

Therefore, we conclude that:

The two variables are associated, i.e., they are not statistically independent.
The Gamma coefficient tells us that the errors in predicting the `dependent variable` are reduced by a proportion of 0.0359 (the value of Gamma coefficient) when you know the `independent variable`.