---
title: "ttest"
author: "Kamila Kolpashnikova"
output:  
      html_document:  
        keep_md: true 
---

# T-test


```{r, echo=FALSE, message=FALSE}
require(knitr)
statapath <- "C:/Program Files (x86)/Stata12/StataSE-64.exe"
opts_chunk$set(engine="stata", engine.path=statapath, comment="")
```

The data here is derived from the [American Time Use Survey](http://www.bls.gov/tus/). For the coding procedures, contact me via email.

## Step 1. Downloading the dataset
You need to download the ATUS_full3.dta file from my github to your working directory.

```{r, results='hide'}
use ATUS_full3.dta, clear
```



## Step 2. Choose a Continuous Variable and Two Groups to Continuous

In the datase, I will use two ordinal variables. One which is time spent on cooking (DVCOOK) and the two groups are represented by the `Female` variable.

```{r, echo=2:3}
use ATUS_full3.dta, clear
sum DVCOOK
tab Female
```

## Step 3. Peform t-test

You use the `ttest` command.

```{r, echo=2}
use ATUS_full3.dta, clear
ttest DVCOOK, by(Female)
```

In the output above, the difference of -27.56011 has a t of -100(-1.0e+02) and it is statistically significant.

## Step 4. Interpretation
The mean time spent on cooking for men is 19.98.The mean time spent on cooking for women is 47.54. This difference is statistically significant, t(159935) = -100, p < .001. In other words, there is a significant difference in time spent on cooking between women and men.
